On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-15 manuscripts/PNAS/review/point_by_point_response.txt

RESPONSE: Yes, thanks for this question. The condition on \ell is perhaps best understood as a condition on \bar m not being too large (given \ell) in order to guarantee the recovery of at least one dictionary element. We have updated the manuscript to reflect this, and the condition now reads “provided \bar m < \ell m / (\ell - 1)” where previously it read “provided p = \bar m − \ell( \bar m − m ) is positive”.

-----------------
-- The paper could also compare the sample complexity results on N with those given in the noise-free case. For example, Table I in reference (18) gives a comparison of sample complexity requirements for different conditions in the noise-free case. Where would the results of this paper stand in that table?

-- Otherwise, all the math seemed correct.
-----------------

RESPONSE: In brief, since all of the theorems and corollaries in [HS15] are corollaries of our theorems, all sample complexities improve by our work. The main point of difference between our statements about sample complexity and those of [HS15] is the assumed constraint on the underlying support set of sparse codes x_i. Our theory of combinatorial designs (the ‘singleton intersection property’) requires only that there be a sufficient number of x_i drawn from each support in a hypergraph satisfying the SIP, whereas the theory in [HS15] requires that a sufficient number of x_i be drawn from *every* support of size k in [m]. Here, we make an explicit comparison with Table I in [HS15] row by row, assuming for our results that sufficient data have been sampled from all supports in a hypergraph H satisfying the SIP:

ROW I. Our result here is |H|(k-1){m \choose k} + 1. We improve on the result in [HS15] by an exponential factor since our theory does not require that data be sampled with every possible support of size k in [m]; as noted in our manuscript, for every k < m there exists a hypergraph H with |H| = m that satisfies the SIP.

ROW 2. Our result here is again |H|(k-1){m \choose k} + 1 with *certainty*, not with *almost certainty* (i.e. probability 1) as in [HS15]. Here the authors in [HS15] have applied probabilistic arguments to achieve an almost certain result with a sample complexity on the same order as that for which we have achieved a certain result by means of our theory of combinatorial designs.

ROW 3. We cannot make a direct comparison here because we have not calculated the probability that a random set of supports satisfy the SIP (see our response to a similar question by the second reviewer). We think this is an interesting problem for the community to solve; regardless, the result of [HS15] here is implied by our more general theory.

ROW 4. Our result here is |H|(k-1){m \choose k} + 1 with probability 1. Here is the only case where the sample complexity stated in [HS15] is technically better than ours, but it differs in flavor: their x_i are assumed to be distributed as (k+1) samples per support for every support of size k in [m], whereas ours supposes that (k-1){m \choose k} codes x_i are distributed over each support in some hypergraph H satisfying the SIP. Regardless, the (technically) lower result of [HS15] is still implied by our more general theory in the case \varepsilon = 0 with the addition of their probabilistic argument.

ROW 5: Again, we cannot make a direct comparison here for the reason stated in ROW 3.

While we appreciate this question, we feel that since our results improve in every case except for one noise-free technicality (ROW 4), and since the results of [HS15] are all entailed anyway by our more general theory, an explicit comparison such as that which we provided above would not be the best use of our limited space in the manuscript; although we are willing to add a discussion about this if necessary. Many thanks for your careful review of the mathematics. We have tried and continue to try to be as clear, concise, and correct as we possibly can to elevate this work into the top echelon of applied mathematical theory papers.

### 2017-08-16 manuscripts/PNAS/chazthm_pnas.tex
poly data

We furthermore note that unlike the requirements of previous works \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}, the matrix $\mathbf{A}$ need not satisfy the spark condition to be recoverable from data. As a simple example for $k=2$, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ where $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$, and take $\mathcal{H}$ to be all consecutive pairs of $[m]$ arranged in cyclic order. Then, $\mathbf{A}$ satisfies the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} but not \eqref{SparkCondition}. %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
This weakening allows for a practical amount of data (polynomial-time for fixed sparsity $k$)  guaranteeing stable dictionaries.

>>>>

We furthermore note that unlike the requirements of previous works \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}, the matrix $\mathbf{A}$ need not satisfy the spark condition to be recoverable from data. As a simple example for $k=2$, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ where $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$, and take $\mathcal{H}$ to be all consecutive pairs of $[m]$ arranged in cyclic order. Then, $\mathbf{A}$ satisfies the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} but not \eqref{SparkCondition}. %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
This weakening allows for a practical (polynomial) amount of data to still guarantee a stable dictionary in the case where the set of sparse supports for the $\mathbf{\bar x}_i$ is known to have a size that grows polynomially in $m, k$.
